---
title: "Effective Memphis Grizzlies Lineups"
author: "Matthew Flaherty"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  github_document: default
editor_options:
  chunk_output_type: console
---

```{r include=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(zoo)
library(nbastatR)
library(future)
library(kableExtra)
library(gt)
```

# Introduction

I have always wanted to look at lineups that a team plays because when certain guys are on the court, the production just looks better than it does when other players are on the court. I am taking advantage of this for this post I will use lineups that the Memphis Grizzlies have played in games to determine which lineups are the most effective in defeating the opponent. 

I measure effectiveness of a lineup in two different ways for this post. The first is effective field goal percent (eFG%) and the second is plus/minus (+/-). eFG%  formula is `(FG + 0.5 * 3PM) / FGA` and adjusts for the fact that a 3-point field goal is worth one more point than a 2-point field goal. For example, suppose Player A goes 4 for 10 with 2 threes, while Player B goes 5 for 10 with 0 threes. Each player would have 10 points from field goals, and thus would have the same effective field goal percentage (50%) ([basketball-reference](https://www.basketball-reference.com/about/glossary.html?lid=header_morelinks)). Thus, it is a better measure of efficiency than field goal percentage. 

+/- "is a measure of the point differential when players are in and out of a game. It is calculated by taking the difference in the score when the player enters the game and subtracting it from the score when the player exits the game. These differences are added up over the entire game to give the score. Thus, it is a measure of the impact a player has on the score of the game when he is in the game without measuring any specific statistic." ([Breakthrough Basketball](https://www.breakthroughbasketball.com/stats/definitions.html)). I can use this for lineups that teams play by getting the initial score for a team when the lineup enters and the score for a team once the lineup on the floor changes. This is a good measurement for how well a lineup plays because higher numbers will result in the team doing well when the lineup is on the court and lower numbers will show that the team did worse with a given lineup on the floor.

# Data

The data that I will be using is from the [nbastatR](https://github.com/abresler/nbastatR) package. I only use data collected for the 2020-2021 NBA season thus far as I want to see which Memphis Grizzlies lineups are playing well together.

This chunk of code is borrowed from [ramirobentes](https://github.com/ramirobentes/NBAblog/blob/master/todo%20pbp%20lineups.R). It is used to get all of the necessary variables that will allow me to manipulate to get only the data for the Grizzlies.

<details>
  <summary>Code</summary>
```{r preprocessing, warning=FALSE, collapse=TRUE}
seconds_to_minutes <- function(x){
  paste0(floor(x / 60), ":", str_pad(x %% 60, side = "left", width = 2, pad = 0))
}

game_logs <- game_logs(seasons = 2021)

games <- game_logs %>%
  select(idGame, slugTeam, slugOpponent, locationGame) %>%
  mutate(slugTeamHome = ifelse(locationGame == "H", slugTeam, slugOpponent),
         slugTeamAway = ifelse(locationGame == "A", slugTeam, slugOpponent)) %>%
  select(-c(slugTeam, slugOpponent, locationGame)) %>%
  distinct(idGame, .keep_all = TRUE)

plan(multiprocess)
play_logs_all <- play_by_play_v2(game_ids = unique(games$idGame))

new_pbp <- play_logs_all %>%
  distinct(idGame, numberEvent, .keep_all = TRUE) %>%   # remove duplicate events
  filter(numberEventMessageType != 18) %>%
  group_by(idGame) %>%
  mutate(numberEvent = row_number()) %>%  # new numberEvent column with events in the right order
  ungroup() %>%
  select(idGame, numberEventMessageType, numberEventActionType, namePlayer1, namePlayer2, namePlayer3,                   
         slugTeamPlayer1, slugTeamPlayer2,  slugTeamPlayer3, numberPeriod, timeQuarter, minuteRemainingQuarter,          
         secondsRemainingQuarter, descriptionPlayHome, numberEvent, descriptionPlayVisitor, scoreHome, scoreAway) %>%
  mutate(shotPtsHome = case_when(
    numberEventMessageType == 3 & !str_detect(descriptionPlayHome, "MISS") ~ 1,                               # Note 1
    numberEventMessageType == 1 & str_detect(descriptionPlayHome, "3PT") ~ 3,                                 # Note 2
    numberEventMessageType == 1 & !str_detect(descriptionPlayHome, "3PT") ~ 2,
    TRUE ~ 0
  )) %>%
  mutate(shotPtsAway = case_when(
    numberEventMessageType == 3 & !str_detect(descriptionPlayVisitor, "MISS") ~ 1,
    numberEventMessageType == 1 & str_detect(descriptionPlayVisitor, "3PT") ~ 3,
    numberEventMessageType == 1 & !str_detect(descriptionPlayVisitor, "3PT") ~ 2,
    TRUE ~ 0
  )) %>%
  group_by(idGame) %>%
  mutate(ptsHome = cumsum(shotPtsHome),
         ptsAway = cumsum(shotPtsAway)) %>%
  ungroup() %>%
  mutate(secsLeftQuarter = (minuteRemainingQuarter * 60) + secondsRemainingQuarter) %>%                       # Note 3 
  mutate(secsStartQuarter = case_when(                                                                        # Note 4
    numberPeriod %in% c(1:5) ~ (numberPeriod - 1) * 720,
    TRUE ~ 2880 + (numberPeriod - 5) * 300
  )) %>%
  mutate(secsPassedQuarter = ifelse(numberPeriod %in% c(1:4), 720 - secsLeftQuarter, 300 - secsLeftQuarter),  # Note 5
         secsPassedGame = secsPassedQuarter + secsStartQuarter) %>%
  left_join(games %>%
              select(idGame, slugTeamHome, slugTeamAway)) %>%
  select(idGame, numberEventMessageType, numberEventActionType, slugTeamHome, slugTeamAway, slugTeamPlayer1, slugTeamPlayer2, 
         slugTeamPlayer3, numberPeriod, timeQuarter, secsPassedGame, numberEvent, namePlayer1, namePlayer2, namePlayer3, 
         descriptionPlayHome, descriptionPlayVisitor, ptsHome, ptsAway, shotPtsHome, shotPtsAway) %>%
  mutate(marginBeforeHome = ptsHome - ptsAway - shotPtsHome + shotPtsAway,
         marginBeforeAway = ptsAway - ptsHome - shotPtsAway + shotPtsHome,
         timeQuarter = str_pad(timeQuarter, width = 5, pad = 0))

subs_made <- new_pbp %>%
  filter(numberEventMessageType == 8) %>%        # Note 6
  mutate(slugTeamLocation = ifelse(slugTeamPlayer1 == slugTeamHome, "Home", "Away")) %>%
  select(idGame, numberPeriod, timeQuarter, secsPassedGame, slugTeamPlayer = slugTeamPlayer1,
         slugTeamLocation, playerOut = namePlayer1, playerIn = namePlayer2) %>%
  pivot_longer(cols = starts_with("player"),
               names_to = "inOut",
               names_prefix = "player",
               values_to = "namePlayer") %>%
  group_by(idGame, numberPeriod, slugTeamPlayer, namePlayer) %>%
  filter(row_number() == 1) %>%
  ungroup()

others_qtr <- new_pbp %>%
  filter(numberEventMessageType != 8) %>%                             
  filter(!(numberEventMessageType == 6 & numberEventActionType %in% c(10, 11, 16, 18, 25))) %>%     # Note 7
  pivot_longer(cols = starts_with("namePlayer"),
               names_to = "playerNumber",
               names_prefix = "namePlayer",
               values_to = "namePlayer") %>%
  mutate(slugTeamPlayer = case_when(playerNumber == 1 ~ slugTeamPlayer1,
                                    playerNumber == 2 ~ slugTeamPlayer2,
                                    playerNumber == 3 ~ slugTeamPlayer3,
                                    TRUE ~ "None")) %>%
  mutate(slugTeamLocation = ifelse(slugTeamPlayer == slugTeamHome, "Home", "Away")) %>%
  filter(!is.na(namePlayer),
         !is.na(slugTeamPlayer)) %>%
  anti_join(subs_made %>%
              select(idGame, numberPeriod, slugTeamPlayer, namePlayer)) %>%    # remove players that were subbed in the quarter
  distinct(idGame, numberPeriod, namePlayer, slugTeamPlayer, slugTeamLocation)

lineups_quarters <- subs_made %>%
  filter(inOut == "Out") %>%
  select(idGame, numberPeriod, slugTeamPlayer, namePlayer, slugTeamLocation) %>%
  bind_rows(others_qtr) %>%
  arrange(idGame, numberPeriod, slugTeamPlayer)

# lineups_quarters %>%
#   count(idGame, numberPeriod, slugTeamPlayer) %>%
#   filter(n != 5)

missing_players_ot <- tribble(
  ~idGame,   ~slugTeamPlayer,          ~namePlayer,     ~numberPeriod,
  22000023,        "DET",           "Delon Wright",           5,
  22000100,        "IND",           "Justin Holiday",         5,
  22000120,        "DEN",           "Gary Harris",            5
) %>%
  left_join(games %>%
              select(idGame, slugTeamHome, slugTeamAway)) %>%
  mutate(slugTeamLocation = ifelse(slugTeamHome == slugTeamPlayer, "Home", "Away")) %>%
  select(-c(slugTeamHome, slugTeamAway))

lineups_quarters <- lineups_quarters %>%
  bind_rows(missing_players_ot) %>%
  arrange(idGame, numberPeriod, slugTeamPlayer)

lineup_subs <- new_pbp %>%
  filter(numberEventMessageType == 8) %>%
  select(idGame, numberPeriod, timeQuarter, secsPassedGame, slugTeamPlayer = slugTeamPlayer1, playerOut = namePlayer1, 
         playerIn = namePlayer2, numberEvent) %>%
  arrange(idGame, numberEvent) %>%
  group_by(idGame, numberPeriod, slugTeamPlayer) %>%
  mutate(row1 = row_number()) %>%
  ungroup() %>%
  left_join(lineups_quarters %>%
              group_by(idGame, numberPeriod, slugTeamPlayer) %>%
              summarise(lineupBefore = paste(sort(unique(namePlayer)), collapse = ", ")) %>%
              ungroup() %>%
              mutate(row1 = 1)) %>%
  select(-row1) %>%
  mutate(lineupBefore = str_split(lineupBefore, ", ")) %>% 
  arrange(idGame, numberEvent) %>%
  group_by(idGame, numberPeriod, slugTeamPlayer) %>%
  mutate(lineupAfter = accumulate2(playerIn, playerOut, ~setdiff(c(..1, ..2), ..3), .init = lineupBefore[[1]])[-1],
         lineupBefore = coalesce(lineupBefore, lag(lineupAfter))) %>%
  ungroup() %>% 
  mutate_all(~map_chr(., ~paste(.x, collapse = ", "))) %>%
  mutate_at(vars("numberEvent", "numberPeriod", "idGame"), ~ as.integer(.)) %>%
  mutate(secsPassedGame = as.numeric(secsPassedGame)) %>%
  arrange(idGame, numberEvent) %>%
  left_join(lineups_quarters %>%
              distinct(idGame, slugTeamPlayer, slugTeamLocation)) %>%
  filter(!is.na(slugTeamLocation))

lineup_game <- new_pbp %>%
  group_by(idGame, numberPeriod) %>%
  mutate(row1 = row_number()) %>%
  ungroup() %>%
  left_join(lineups_quarters %>%
              group_by(idGame, numberPeriod, slugTeamLocation) %>%
              summarise(lineupBefore = paste(sort(unique(namePlayer)), collapse = ", ")) %>%
              ungroup() %>%
              pivot_wider(names_from = slugTeamLocation,
                          names_prefix = "lineupInitial",
                          values_from = lineupBefore) %>%
              mutate(row1 = 1)) %>%
  select(-row1) %>%
  left_join(lineup_subs %>%
              mutate(lineupBeforeHome = ifelse(slugTeamLocation == "Home", lineupBefore, NA),
                     lineupAfterHome = ifelse(slugTeamLocation == "Home", lineupAfter, NA),
                     lineupBeforeAway = ifelse(slugTeamLocation == "Away", lineupBefore, NA),
                     lineupAfterAway = ifelse(slugTeamLocation == "Away", lineupAfter, NA)) %>%
              select(idGame, numberPeriod, timeQuarter, secsPassedGame, numberEvent, slugTeamPlayer1 = slugTeamPlayer,
                     contains("Home"), contains("Away"))) %>%
  mutate_at(vars(c(lineupBeforeHome, lineupAfterHome)), ~ ifelse(!is.na(lineupInitialHome), lineupInitialHome, .)) %>%
  mutate_at(vars(c(lineupBeforeAway, lineupAfterAway)), ~ ifelse(!is.na(lineupInitialAway), lineupInitialAway, .)) %>%
  select(-starts_with("lineupInitial")) %>%
  group_by(idGame, numberPeriod) %>%
  mutate(lineupHome = na.locf(lineupAfterHome, na.rm = FALSE),
         lineupAway = na.locf(lineupAfterAway, na.rm = FALSE),
         lineupHome = ifelse(is.na(lineupHome), na.locf(lineupBeforeHome, fromLast = TRUE, na.rm = FALSE), lineupHome),
         lineupAway = ifelse(is.na(lineupAway), na.locf(lineupBeforeAway, fromLast = TRUE, na.rm = FALSE), lineupAway),
         lineupHome = str_split(lineupHome, ", "),
         lineupAway = str_split(lineupAway, ", "),
         lineupHome = map_chr(lineupHome, ~ paste(sort(.), collapse = ", ")),
         lineupAway = map_chr(lineupAway, ~ paste(sort(.), collapse = ", "))) %>%
  ungroup() %>%
  select(-c(starts_with("lineupBefore"), starts_with("lineupAfter")))


lineup_game_stats <- lineup_game %>%
  mutate(canSub = case_when(numberEventMessageType == 5 & !numberEventActionType %in% c(1, 2) ~ 1,    # dead ball turnovers
                            numberEventMessageType == 6 & numberEventActionType != 16 ~ 1,            # fouls
                            numberEventMessageType == 11 & numberEventActionType != 4 ~ 1,
                            numberEventMessageType == 7 & numberEventActionType == 5 ~ 1,             # kickballs
                            numberEventMessageType == 4 & numberEventActionType == 0 & !str_detect(str_to_upper(descriptionPlayHome), "OFF:") ~ 1,
                            numberEventMessageType == 4 & numberEventActionType == 0 & !str_detect(str_to_upper(descriptionPlayVisitor), "OFF:") ~ 1,
                            TRUE ~ 0)) %>%
  mutate(secsPassedGame2 = ifelse(timeQuarter == "12:00" &
                                    (str_detect(str_to_lower(descriptionPlayHome), "technical") |
                                       str_detect(str_to_lower(descriptionPlayVisitor), "technical")),
                                  secsPassedGame + 0.5, secsPassedGame)) %>%    # Note 4
  group_by(idGame, numberPeriod, secsPassedGame) %>%
  mutate(numberNew = ifelse(numberEventMessageType == 3 & numberEventActionType == 12, 
                            paste(numberEvent[numberEventMessageType == 3 & numberEventActionType == 11], collapse = ", "), 
                            as.character(numberEvent)),
         numberNew = ifelse(numberEventMessageType == 3 & numberEventActionType %in% c(14, 15), 
                            paste(numberEvent[numberEventMessageType == 3 & numberEventActionType == 13], collapse = ", "),
                            numberNew)) %>%
  mutate(numberNew = str_split(numberNew, ", "),
         numberNew = map(numberNew, ~as.numeric(.)),
         numberNew = map2_dbl(numberNew, numberEvent, ~ max(.x[.x <= .y]))) %>%
  ungroup() %>%
  arrange(idGame, numberNew, numberEvent) %>%
  group_by(idGame) %>%
  mutate(newptsHome = cumsum(shotPtsHome),
         newptsAway = cumsum(shotPtsAway)) %>%
  group_by(idGame, numberPeriod, secsPassedGame2) %>%
  mutate(subOpp = cumsum(canSub)) %>%
  group_by(idGame = as.character(idGame), numberPeriod = as.character(numberPeriod), subOpp, secsPassedGame2 = as.character(secsPassedGame2)) %>%
  mutate(hasFouls = sum(numberEventMessageType == 3)) %>%
  mutate(newptsHome = ifelse(hasFouls > 0,
                             newptsHome[row_number() == max(row_number()[numberEventMessageType == 3])],
                             newptsHome),
         newptsAway = ifelse(hasFouls > 0,
                             newptsAway[row_number() == max(row_number()[numberEventMessageType == 3])],
                             newptsAway)) %>%
  ungroup() %>%
  select(-hasFouls) %>%
  select(-c(numberNew, secsPassedGame2)) %>%
  mutate_all(~ as.character(.)) %>%
  mutate(secsPassedGame = as.numeric(secsPassedGame),
         numberEvent = as.numeric(numberEvent))

lineup_stats <- lineup_game_stats %>%
  select(idGame, numberEvent, slugTeamHome, slugTeamAway, numberPeriod, timeQuarter, secsPassedGame, 
         newptsHome, newptsAway, lineupHome, lineupAway) %>%
  pivot_longer(cols = starts_with("lineup"),
               names_to = "lineupLocation",
               names_prefix = "lineup",
               values_to = "lineup") %>%
  mutate(ptsTeam = ifelse(lineupLocation == "Home", newptsHome, newptsAway),
         ptsOpp = ifelse(lineupLocation == "Away", newptsHome, newptsAway),
         slugTeam = ifelse(lineupLocation == "Home", slugTeamHome, slugTeamAway),
         slugOpp = ifelse(lineupLocation == "Away", slugTeamHome, slugTeamAway)) %>%
  distinct(idGame, slugTeam, slugOpp, numberPeriod, timeQuarter, secsPassedGame, ptsTeam, ptsOpp, lineup, 
           teamLocation = lineupLocation, numberEvent) %>%
  arrange(idGame, numberEvent) %>%
  group_by(idGame, slugTeam) %>%
  mutate(lineupChange = lineup != lag(lineup),
         lineupChange = coalesce(lineupChange, FALSE)) %>%
  group_by(idGame, slugTeam) %>%
  mutate(lineupStint = cumsum(lineupChange)) %>%
  ungroup() %>%
  arrange(idGame, lineupStint, numberEvent) %>%
  group_by(idGame, slugTeam, lineup, lineupStint) %>%
  summarise(initialScoreTeam = ptsTeam[row_number() == min(row_number())],
            initialScoreOpp = ptsOpp[row_number() == min(row_number())],
            finalScoreTeam = ptsTeam[row_number() == max(row_number())],
            finalScoreOpp =  ptsOpp[row_number() == max(row_number())],
            initialTime = secsPassedGame[row_number() == min(row_number())],
            finalTime = secsPassedGame[row_number() == max(row_number())]) %>%
  ungroup() %>%
  arrange(idGame, lineupStint) %>%
  group_by(idGame, slugTeam) %>%                              
  mutate(finalTime = ifelse(row_number() == max(row_number()), finalTime, lead(initialTime))) %>%  
  ungroup() %>%
  mutate(across(c(contains("Score")), ~ as.numeric(.), .names = "{col}")) %>%
  mutate(totalScoreTeam = finalScoreTeam - initialScoreTeam,
         totalScoreOpp = finalScoreOpp - initialScoreOpp,
         netScoreTeam = totalScoreTeam - totalScoreOpp,
         totalTime = finalTime - initialTime) %>%
  arrange(idGame, lineupStint)

# indiv_stats <- lineup_stats %>%
#   separate_rows(lineup, sep = ", ") %>%
#   group_by(namePlayer = lineup, idGame, slugTeam) %>%
#   summarise(totalPlusMinus = sum(netScoreTeam),
#             totalSecs = sum(totalTime)) %>%
#   ungroup() %>%
#   arrange(-totalPlusMinus)
# 
# indiv_stats %>%
#   group_by(namePlayer) %>%
#   summarise(seasonPM = sum(totalPlusMinus),
#             seasonSecs = sum(totalSecs)) %>%
#   ungroup() %>%
#   arrange(-seasonPM) %>%
#   mutate(seasonMin = paste0(floor(seasonSecs / 60), ":", str_pad(round(seasonSecs %% 60, 0), side = "left", width = 2, pad = 0))) %>%
#   select(-seasonSecs)
# 
# lineup_stats %>%
#   group_by(lineup) %>%
#   summarise(seasonPM = sum(netScoreTeam),
#             seasonSecs = sum(totalTime)) %>%
#   ungroup() %>%
#   arrange(-seasonPM, -seasonSecs) %>%
#   mutate(seasonMin = paste0(floor(seasonSecs / 60), ":", str_pad(round(seasonSecs %% 60, 0), side = "left", width = 2, pad = 0))) %>%
#   select(-seasonSecs)

rm(lineup_game, games, df_nba_player_dict, lineup_subs, missing_players_ot, others_qtr, subs_made)

```
</details>

# Effective Lineups

<details>
  <summary>Code</summary> 
```{r stats_2021}
game_filter <- game_logs %>%
  filter(yearSeason == 2021)%>%
  summarise(max_game = max(idGame)) %>%
  pull(max_game)
```
</details>

The lineups in this table are the top 15 lineups that the Grizzlies have used. I was not sure what the cutoff for minutes played by a lineup should be so I took the top 15 lineups for minutes played.

```{r top_15_used_lineups}
grizz_lineups <- lineup_stats %>%
  filter(idGame <= game_filter) %>%
  filter(slugTeam == "MEM") %>%
  group_by(lineup) %>%
  summarise(min_played = round(sum(totalTime)/60,3),
            bpm = round(sum(netScoreTeam)/min_played,3)) %>%
  ungroup() %>%
  arrange(-min_played) %>%
  slice(1:15)

grizz_lineups %>%
  mutate(Rank = row_number()) %>%
  rename(Lineup = lineup,
         'Minutes Played' = min_played,
         "PM/min" = bpm) %>%
  select(Rank, everything())%>%
  gt() %>%
  tab_header(
    title = "Memphis Grizzlies Most Popular Lineups",
    subtitle = "2021 Season"
  ) %>%
  data_color(
    columns = vars("Minutes Played"),
    colors = "yellow")
```

The lineup that has the most minutes played is not the typical starting lineup for the Grizzlies as they are not at full health. Ja Morant had to miss a few weeks with an ankle injury and Jaren Jackson Jr. has yet to dress out for a game. Thus, the lineup with the most minutes played was the starting group while Morant was out.

## Effective FG %

I am going to plot the eFG% vs teams wins in a season to see if there is any correlation. If there is, then the Grizzlies should look to maximize the time that the lineups with good eFG% have on the court.

<details>
  <summary>Code</summary>
```{r collapse=TRUE, warning=FALSE}
team_stats <- bref_teams_stats(seasons = 2016:2020)
team_stats <- as.data.frame(team_stats$dataTable)

team_efg <- team_stats %>%
  mutate(efg = (team_stats$fgmPerGameTeam + 0.5 * team_stats$fg3mPerGameTeam)/ team_stats$fgaPerGameTeam)
```
</details>
 
```{r efg_vs_wins}
ggplot(team_efg, aes(x = winsTeam, y = efg))+
  geom_point()+
  annotate(geom = "text", x = 65, y = 0.53, label = paste("r = ", round(cor(team_efg$winsTeam,team_efg$efg),3)), col = "blue") +
  geom_smooth(method = "lm") +
  labs(
    title = "Effective FG % vs Wins",
    subtitle = "Seasons 2016-2020",
    x = "Wins",
    y = "Effective Field Goal %",
    caption = "Data from nbastatR\nPlot by @flahermi"
  ) +
  theme_bw() +
  theme(
    plot.subtitle = element_text(size = 8, colour = "grey70"),
    plot.title = element_text(size = 18, face = "bold")
  ) 
```

There is moderate correlation between eFG% and wins. Thus, this statistic could be used as a measurement of how well a lineup is playing. Using the top 15 lineups played, I can find the most efficient lineup using eFG%. 

<details>
  <summary>Code</summary>
```{r efg}
grizz_best <- lineup_game_stats %>%
  filter(idGame <= game_filter) %>%
  filter(numberEventMessageType %in% c(1,2)) %>%
  mutate(shotPts = if_else(slugTeamPlayer1 == slugTeamHome, shotPtsHome, shotPtsAway),
         lineup = if_else(slugTeamPlayer1 == slugTeamHome, lineupHome, lineupAway)) %>%
  count(lineup, shotPts) %>%
  filter(lineup %in% grizz_lineups$lineup) %>%
  pivot_wider(names_from = shotPts,
              values_from = n,
              values_fill = 0) %>%
  janitor::clean_names()%>%
  mutate(effective_fg = round(((x2 + x3) + 0.5 * x3) / (x0 + x2 + x3),4)*100) %>%
  arrange(-effective_fg)

#  add minutes played and BPM by this lineup
#  this will allow me to assess if this lineup needs to play more
grizz_best <- grizz_best %>%
  mutate(min_played = NA,
         bpm = NA)

for (i in 1:length(grizz_best$lineup)) {
  for (j in 1:length(grizz_lineups$lineup)) {
    if (grizz_best$lineup[i] == grizz_lineups$lineup[j]){
      grizz_best$min_played[i] <- grizz_lineups$min_played[j]
      grizz_best$bpm[i] <- grizz_lineups$bpm[j]
    } 
  }
}
```
</details>

```{r}
grizz_best %>%
  mutate(Rank = row_number(),
         FG = x0 + x2 + x3,
         "2FGM" = x2,
         "3FGM" = x3) %>%
  rename(Lineup = lineup,
         'Minutes Played' = min_played,
         'PM/min' = bpm, 
         'Effective FG%' = effective_fg) %>%
  select(Rank, Lineup, FG, "2FGM", "3FGM", "Effective FG%", "Minutes Played", 'PM/min',-x0, -x2, -x3) %>%
  gt() %>%
  tab_header(
    title = "Memphis Grizzlies Lineups by eFG%",
    subtitle = "2021 Season"
  ) %>%
  data_color(
    columns = vars("Effective FG%"),
    colors = "yellow")
```

I find it interesting that Clarke, Bane, and Jones are in each of the top 3 lineup for this statistic. At the same time, these are players who do not take head-scratching shots and they leave the viewer confident that the shot that they have chosen was a good shot. The second ranking lineup in this category probably has a good eFG% because it consists of players who can hit shots. For example, Dieng plays the five; however, the Grizzlies will use him to set a screen and stay on the perimeter showing that they trust his three-point shooting ability. Brandon Clarke is also a player who has the height of a four or five but rarely posts up and will be seen often on the perimeter.

## Plus/Minus

+/- can be graphed from the team perspective as well. Any correlation between +/- and the number of wins a team has can signify that the variable may be used to judge how a team is doing. I attempt to find correlation between +/- and team wins in the graph below.

```{r pm_wins}
team_pm <- team_stats %>%
  mutate(pm = ptsTotalsTeam - ptsTotalsOpponent)
  
ggplot(team_pm, aes(x = winsTeam, y = pm))+
  geom_point()+
  annotate(geom = "text", x = 65, y = 500, label = paste("r = ", round(cor(team_pm$winsTeam,team_pm$pm),3)), col = "blue") +
  geom_smooth(method = "lm") +
  labs(
    title = "Team Plus/Minus vs Wins",
    subtitle = "Seasons 2016-2020",
    x = "Wins",
    y = "Plus/Minus",
    caption = "Data from nbastatR\nPlot by @flahermi"
  ) +
  theme_bw() +
  theme(
    plot.subtitle = element_text(size = 8, colour = "grey70"),
    plot.title = element_text(size = 18, face = "bold")
  ) 
```

+/- has strong positive correlation with number of wins in a season so this provides evidence that teams should strive to maximize their +/- rating. One way to do this is to play lineups that have a good +/-. I want to look organize the table by +/- rating now because I think that the Grizzlies should give more minutes to the lineups that have good +/- rating.

```{r bpm}
grizz_best %>%
  arrange(-bpm) %>%
  mutate(Rank = row_number(),
         FG = x0 + x2 + x3,
         "2FGM" = x2,
         "3FGM" = x3) %>%
  rename(Lineup = lineup,
         'Minutes Played' = min_played,
         'PM/min' = bpm, 
         'Effective FG%' = effective_fg) %>%
  select(Rank, Lineup, FG, "2FGM", "3FGM", "Effective FG%", "Minutes Played", 'PM/min',-x0, -x2, -x3) %>%
  gt() %>%
  tab_header(
    title = "Memphis Grizzlies Lineups by +/-",
    subtitle = "2021 Season"
  ) %>%
  data_color(
    columns = vars("PM/min"),
    colors = "yellow")
```

I think that the best-ranking lineup in this statistic may need to be taken with a grain of salt because the Grizzlies best-ranking lineup does not contain a single starter. This means that the other team probably does not have their starters in either. This is not all bad new as it likely means that the Grizzlies are gaining the advantage when both teams have their bench players in the game; however, any person on the Grizzlies coaching staff should **NOT** be pushing to start this lineup until they have proved themselves against other teams' starters. The third best lineup for +/- consists of mainly starters and still comes out better than its opponents so it's safe to say that the Grizzlies' starters can help the team gain an advantage as well as the bench players. 

# Conclusion

Clarke, Bane, and T. Jones are playing effective basketball together which is good news for the Grizzlies because these are role players who mainly start the game on the bench when the whole team is healthy. Effective bench players are a key for success because the coach needs to trust that these players will not enter a game and do anything that may hinder the team's chance of winning. Especially with the limited time that they get, they must make the most of it. I think that these three players should continue to get playing time because eFG% is positively correlated to wins and when these players are on the court, the lineup useually has a good eFG%.

The Grizzlies' bench players are outperforming opponents' bench players which is crucial because it shows that the bench players are coming into each game and putting the Grizzlies in a better situation than the team was in before they entered the game. The coaching staff must realize that these players' +/- are good because they are playing against opponents with their skill level so if they get moved into the starting position, the +/- rating may decrease a little. Having said that, if the Grizzlies are going to have a consistent bench lineup that plays together, then I think Melton, Bane, Dieng, Allen, and Tillman would put the Grizzlies in the best situation. This lineup has the best +/- on the team and +/- has been shown to have a strong positive correlation with wins. 



