---
title: "Data Visualization"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(readr)

fantasy <- read_csv("data/fantasy.csv")
```

```{r bpm_fpts}
ggplot(fantasy, aes(bpm, fpts_100)) +
  geom_point() +
  annotate(geom = "text", x = -20, y = -25, label = paste("r = ",format(round(cor(fantasy$bpm, fantasy$fpts_100),4))), colour = "blue") +
  labs(
    title = "Fantasy Points vs BPM",
    x = "BPM",
    y = "Fantasy Points"
  ) +
  theme_classic() +
  geom_smooth(method = "lm") 
```

```{r per_fpts}
ggplot(fantasy, aes(per, fpts_100)) +
  geom_point()+
  annotate(geom = "text", x = -5, y = -10, label = paste("r = ",format(round(cor(fantasy$per, fantasy$fpts_100),4))), colour = "blue") +
  labs(
    title = "Fantasy Points vs PER",
    x = "PER",
    y = "Fantasy Points"
  ) +
  theme_classic() +
  geom_smooth(method = "lm") 
  

```

```{r obpm_fpts}
ggplot(fantasy, aes(obpm, fpts_100)) +
  geom_point()+
  annotate(geom = "text", x = -5, y = -10, label = paste("r = ",format(round(cor(fantasy$obpm, fantasy$fpts_100),4))), colour = "blue") +
  labs(
    title = "Fantasy Points vs OBPM",
    x = "OBPM",
    y = "Fantasy Points"
  ) +
  theme_classic() +
  geom_smooth(method = "lm") 
  
```

```{r pts_fpts}
ggplot(fantasy, aes(pts, fpts_100)) +
  geom_point()+
  annotate(geom = "text", x = 45, y = 30, label = paste("r = ",format(round(cor(fantasy$pts, fantasy$fpts_100),4))), colour = "blue") +
  labs(
    title = "Fantasy Points vs Points",
    x = "Points",
    y = "Fantasy Points"
  ) +
  theme_classic() +
  geom_smooth(method = "lm") 
  
```

# BPM + PER

BPM and PER both have a strong correlation with fantasy points so I add the two statistics together and test the correlation. If it is a stronger correlation than PER (`r format(round(cor(fantasy$per, fantasy$fpts_100),4))`), then I will use BPM+PER to determine which players to pickup.

```{r bpm_per}
fantasy <- fantasy %>%
  mutate(bpm_per = bpm +per)

ggplot(fantasy, aes(bpm_per, fpts_100)) +
  geom_point() +
  annotate(geom = "text", x = 0, y = 50, label = paste("r = ",format(round(cor(fantasy$bpm_per, fantasy$fpts_100),4))), colour = "blue") +
  labs(
    title = "Fantasy Points vs BPM+PER",
    x = "BPM+PER",
    y = "Fantasy Points"
  ) +
  theme_classic() +
  geom_smooth(method = "lm") 
```

BPM+PER has a correlation of `r format(round(cor(fantasy$bpm_per, fantasy$fpts_100),4))`. This is about the same as the correlation PER has with fantasy points but still less so I will just use PER instead of BPM+PER

## Multiply by coefficient

What if I multiply the variables by their correlation coefficient when measured individually against `fpts`?

```{r}
fantasy <- fantasy %>%
  mutate(measure = (0.8362*bpm) + (0.9206*per))

ggplot(fantasy, aes(measure, fpts_100)) +
  geom_point() +
  annotate(geom = "text", x = 0, y = 50, label = paste("r = ",format(round(cor(fantasy$measure, fantasy$fpts_100),4))), colour = "blue") +
  labs(
    title = "Fantasy Points vs BPM+PER",
    x = "BPM+PER",
    y = "Fantasy Points"
  ) +
  theme_classic() +
  geom_smooth(method = "lm") 
```

I get a correlation coefficient of `r format(round(cor(fantasy$measure, fantasy$fpts_100),4))`. Still less than that of `per` and `fpts`; therefore, I still have reason to use `per` alone.

# Linear Regression

Now I will get correlation coefficients using a linear regression to multiply the statistics by.

## Scale the data

It is important to scale the data so the variables are within the same range and neither variable can outweigh the other.

```{r}
fantasy$bpm <- scale(fantasy$bpm)

fantasy$per <- scale(fantasy$per)

fantasy$fpts_100 <- scale(fantasy$fpts_100)
```

## Linear Regression Model

Now that I have scaled my data, I can use a linear regression model to get correlation coefficients for the variables. I will take these coefficients and multiply the respective variable by the coefficient.

```{r}
lm(data = fantasy, fpts_100~bpm + per)

fantasy <- fantasy %>%
  mutate(coeff = (7.400e-02*bpm) + (8.546e-01*per))

ggplot(fantasy, aes(coeff, fpts_100)) +
  geom_point() +
  annotate(geom = "text", x = -2, y = 5, label = paste("r = ",format(round(cor(fantasy$coeff, fantasy$fpts_100),4))), colour = "blue") +
  labs(
    title = "Fantasy Points vs BPM+PER",
    x = "BPM+PER",
    y = "Fantasy Points"
  ) +
  theme_classic() +
  geom_smooth(method = "lm") 
```

Doing this gives me a correlation of `rformat(round(cor(fantasy$coeff, fantasy$fpts_100),4))` which is greater than `per` but not by a significant amount. 

Taking the mean of the two variables and its respective correlation coefficients to if that has any impact.

```{r}
fantasy <- fantasy %>%
  mutate(mean_bpm_per = ((7.400e-02*bpm) + (8.546e-01*per))/2)

ggplot(fantasy, aes(mean_bpm_per, fpts_100)) +
  geom_point() +
  annotate(geom = "text", x = -2, y = 5, label = paste("r = ",format(round(cor(fantasy$mean_bpm_per, fantasy$fpts_100),4))), colour = "blue") +
  labs(
    title = "Fantasy Points vs BPM+PER",
    x = "BPM+PER",
    y = "Fantasy Points"
  ) +
  theme_classic() +
  geom_smooth(method = "lm") 
```

I can see that it does not have any impact.
