---
title: "Import Data"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Problem

In basketball and many other sports, the goal of the game is to score more points than the other team. One way to do this is to limit the other team's possessions. This could potentially limit your team's possessions. Thus your possessions to the other team's possessions will be 1:1. Offensive rebounds are a way to increase this ratio in your favor, 2:1,3:1,...,N:1. What if we were able to determine the chance of us getting an offensive rebound? Then we could use the necessary tactics to increase our chances of getting an offensive rebound if they were low.

# Libraries

```{r}
library(tidyverse)
library(assertr)
library(h2o)
library(glue)
library(lemon)
knit_print.data.frame <- lemon_print
library(knitr)
h2o.init(nthreads = -1, #Number of threads -1 means use all cores on your machine
         max_mem_size = "8G")  #max mem size is the maximum memory to allocate to H2O
```

# Read train data pbp

```{r}
training_data_pbp <- read_csv("training_data_pbp.csv", 
    col_types = cols(X1 = col_skip(), away_team_id = col_character(), 
        def_team_id = col_character(), game_id = col_character(), 
        home_team_id = col_character(), off_team_id = col_character(), 
        playbyplayorder_id = col_character(), 
        playerid_off_player_1 = col_character(), 
        playerid_off_player_2 = col_character(), 
        playerid_off_player_3 = col_character(), 
        playerid_off_player_4 = col_character(),
        playerid_off_player_5 = col_character(),
        playerid_def_player_1 = col_character(),
        playerid_def_player_2 = col_character(),
        playerid_def_player_3 = col_character(),
        playerid_def_player_4 = col_character(),
        playerid_def_player_5 = col_character(),
        reb_player_id = col_character(), 
        shooter_player_id = col_character()))
```

# train pbp description

name - description
game_id - ID for the game
playbyplayorder_id - ID for the play by play event in this row
row_type - Type of play by play event, shot or final free throw (e.g. the last free throw taken during a trip to the line that could be rebounded if missed)
F.OREB - Whether a shot was offensive rebounded (Yes or No)
REBOFFENSIVE - Whether a shot was offensive rebounded (0 = no, 1 = yes)
REBDEFFENSIVE - Whether a shot was defensive rebounded (0 = no, 1 = yes)
OFF_TEAM_ID - Team ID of the team on offense
DEF_TEAM_ID - Team ID of the team on defense
HOME_TEAM_ID - Team ID of the home team
AWAY_TEAM_ID - Team ID of the away team
FTMADE - Did this play result in a made free throw? (0 = no, 1 = yes)
FTMISSED - Did this play result in a missed free throw? (0 = no, 1 = yes)
FG2MADE - Did this play result in a made 2 point attempt? (0 = no, 1 = yes)
FG2MISSED - Did this play result in a missed 2 point attempt? (0 = no, 1 = yes)
FG2ATTEMPTED - Was a two point shot attempted on this play? (0 = no, 1 = yes)
FG3MADE - Did this play result in a made 3 point attempt? (0 = no, 1 = yes)
FG3MISSED - Did this play result in a missed 3 point attempt? (0 = no, 1 = yes)
FG3ATTEMPTED - Was a three point shot attempted on this play? (0 = no, 1 = yes)
SECONDCHANCE - Indicator for a second chance opportunity. Increments for each subsequent chance following another offensive rebound. (e.g. if there are 3 shots on a possession, the first will have 0, the second will have 1,the third will have 2)
ANDONE - Was this play a shooting foul on made shot? (0 = no, 1 = yes)
INBONUS - Is the offensive team in the bonus? (0 = no, 1 = yes)
EVENTDESCRIPTION - A description of the play by play event 
ACTIONDESCRIPTION - A more specific description of the play by play event
SHOOTER_PLAYER_ID - The player_id of the shooter
REB_PLAYER_ID - The player_id of the rebounder
PLAYER_ID_OFF_PLAYER_1 - The ID of one of the 5 offensive players on the court (no particular order)
PLAYER_ID_OFF_PLAYER_2 - The ID of one of the 5 offensive players on the court (no particular order)
PLAYER_ID_OFF_PLAYER_3 - The ID of one of the 5 offensive players on the court (no particular order)
PLAYER_ID_OFF_PLAYER_4 - The ID of one of the 5 offensive players on the court (no particular order)
PLAYER_ID_OFF_PLAYER_5 - The ID of one of the 5 offensive players on the court (no particular order)
PLAYER_ID_DEF_PLAYER_1 - The ID of one of the 5 defensive players on the court (no particular order)
PLAYER_ID_DEF_PLAYER_2 - The ID of one of the 5 defensive players on the court (no particular order)
PLAYER_ID_DEF_PLAYER_3 - The ID of one of the 5 defensive players on the court (no particular order)
PLAYER_ID_DEF_PLAYER_4 - The ID of one of the 5 defensive players on the court (no particular order)
PLAYER_ID_DEF_PLAYER_5 - The ID of one of the 5 defensive players on the court (no particular
order)

# Read train data loc

```{r}
training_data_loc <- read_csv("training_data_loc.csv", 
    col_types = cols(X1 = col_skip(), game_id = col_character(), 
        playbyplayorder_id = col_character()))
```

name - description
GAME_ID - ID for the game
PLAYBYPLAYORDER_ID - ID for the play by play event in this row.
ROW_TYPE - Type of play by play event, shot or final free throw (e.g. the last free throw taken during a trip to the line that could be rebounded if missed)
ATSHOT_LOC_X_OFF_PLAYER_1 - The x coordinate of the location of Offensive Player 1 at the time of the shot
ATSHOT_LOC_Y_OFF_PLAYER_1 - The y coordinate of the location of Offensive Player 1 at the time of the shot
ATRIM_LOC_X_OFF_PLAYER_1 - The x coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_OFF_PLAYER_1 - The y coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_OFF_PLAYER_2 - The x coordinate of the location of Offensive Player 1 at the time of the shot
ATSHOT_LOC_Y_OFF_PLAYER_2 - The y coordinate of the location of Offensive Player 1 at the time of the shot
ATRIM_LOC_X_OFF_PLAYER_2 - The x coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_OFF_PLAYER_2 - The y coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_OFF_PLAYER_3 - The x coordinate of the location of Offensive Player 1 at the time of the shot
ATSHOT_LOC_Y_OFF_PLAYER_3 - The y coordinate of the location of Offensive Player 1 at the time of the shot
ATRIM_LOC_X_OFF_PLAYER_3 - The x coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_OFF_PLAYER_3 - The y coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_OFF_PLAYER_4 - The x coordinate of the location of Offensive Player 1 at the time of the shot
ATSHOT_LOC_Y_OFF_PLAYER_4 - The y coordinate of the location of Offensive Player 1 at the time of the shot
ATRIM_LOC_X_OFF_PLAYER_4 - The x coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_OFF_PLAYER_4 - The y coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_OFF_PLAYER_5 - The x coordinate of the location of Offensive Player 1 at the time of the shot
ATSHOT_LOC_Y_OFF_PLAYER_5 - The y coordinate of the location of Offensive Player 1 at the time of the shot
ATRIM_LOC_X_OFF_PLAYER_5 - The x coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_OFF_PLAYER_5 - The y coordinate of the location of Offensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_DEF_PLAYER_1 - The x coordinate of the location of defensive Player 1 at the time of the shot
ATSHOT_LOC_Y_DEF_PLAYER_1 - The y coordinate of the location of defensive Player 1 at the time of the shot
ATRIM_LOC_X_DEF_PLAYER_1 - The x coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_DEF_PLAYER_1 - The y coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_DEF_PLAYER_2 - The x coordinate of the location of defensive Player 1 at the time of the shot
ATSHOT_LOC_Y_DEF_PLAYER_2 - The y coordinate of the location of defensive Player 1 at the time of the shot
ATRIM_LOC_X_DEF_PLAYER_2 - The x coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_DEF_PLAYER_2 - The y coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_DEF_PLAYER_3 - The x coordinate of the location of defensive Player 1 at the time of the shot
ATSHOT_LOC_Y_DEF_PLAYER_3 - The y coordinate of the location of defensive Player 1 at the time of the shot
ATRIM_LOC_X_DEF_PLAYER_3 - The x coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_DEF_PLAYER_3 - The y coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_DEF_PLAYER_4 - The x coordinate of the location of defensive Player 1 at the time of the shot
ATSHOT_LOC_Y_DEF_PLAYER_4 - The y coordinate of the location of defensive Player 1 at the time of the shot
ATRIM_LOC_X_DEF_PLAYER_4 - The x coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_DEF_PLAYER_4 - The y coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATSHOT_LOC_X_DEF_PLAYER_5 - The x coordinate of the location of defensive Player 1 at the time of the shot
ATSHOT_LOC_Y_DEF_PLAYER_5 - The y coordinate of the location of defensive Player 1 at the time of the shot
ATRIM_LOC_X_DEF_PLAYER_5 - The x coordinate of the location of defensive Player 1 at the time the ball hit the rim
ATRIM_LOC_Y_DEF_PLAYER_5 - The y coordinate of the location of defensive Player 1 at the time the ball hit the rim

# Player Position Data

```{r}
player_pos_data <- read_csv("player_pos_data.csv", 
    col_types = cols(X1 = col_character(), 
        player_id = col_character(), position = col_factor(levels = c("1", 
            "2", "3", "4", "5")), position2 = col_factor(levels = c("1", 
            "2", "3", "4", "5"))))%>%
  select(-X1)
```

# Player position description

name - description
PLAYER_ID - An ID identifying the player
POSITION - The player's primary position
POSITION2 - The player's secondary position (if applicable)

# Player Rebounding Totals

```{r}
player_reb_data <- read_csv("player_reb_data.csv", 
    col_types = cols(X1 = col_skip(), def_rebchances = col_integer(), 
        def_rebs = col_integer(), games = col_integer(), 
        off_rebchances = col_integer(), off_rebs = col_integer(), 
        player_id = col_character()))
```

## Player rebounding description

name - description
PLAYER_ID - An ID identifying the player
GAMES - The number of games this player played in the games that are included in both the training and testing data.
MINUTES -The number of minutes this player played in the games that are included in both the training and testing data.
OFF_REBOUNDS - The number of offensive rebounds this player grabbed in the games that are included in both the training and testing data.
DEF_REBOUNDS - The number of defensive rebounds this player grabbed in the games that are included in both the training and testing data.
OFF_REBCHANCES - The number of chances this player had to secure an offensive rebound during the games that are included in both the training and testing data. Note that a player simply needs to be on offense for a missed shot to record an "off_rebchance". This variable does not indicate, e.g., whether the player was in position to potentially secure a rebound.
DEF_REBCHANCES - The number of chances this player had to secure a defensive rebound during the games that are included in both the training and testing data. Note that a player simply needs to be on defense for a missed shot to record a "def_rebchance". This variable does not indicate, e.g., whether the player was in position to potentially secure a rebound.

# Load testing pbp data

```{r}
testing_data_pbp <- read_csv("testing_data_pbp.csv", 
    col_types = cols(X1 = col_skip(), away_team_id = col_character(), 
        def_team_id = col_character(), game_id = col_character(), 
        home_team_id = col_character(), off_team_id = col_character(), 
        playbyplayorder_id = col_character(), 
        shooter_player_id = col_character(),
        playbyplayorder_id = col_character(), 
        playerid_off_player_1 = col_character(), 
        playerid_off_player_2 = col_character(), 
        playerid_off_player_3 = col_character(), 
        playerid_off_player_4 = col_character(),
        playerid_off_player_5 = col_character(),
        playerid_def_player_1 = col_character(),
        playerid_def_player_2 = col_character(),
        playerid_def_player_3 = col_character(),
        playerid_def_player_4 = col_character(),
        playerid_def_player_5 = col_character()))
```

# Load testing location data
```{r}
testing_data_loc <- read_csv("testing_data_loc.csv", 
    col_types = cols(X1 = col_skip(), game_id = col_character(),
                     playbyplayorder_id = col_character()))
```

